### 2019-09-01 paper/paper.tex
Edited down to results.

\documentclass{bmd2019a}

\usepackage{siunitx}
\usepackage{amsmath}
\usepackage{booktabs}

\begin{document}

\begin{flushleft}
{\fontsize{16pt}{20pt}\selectfont%
  Expanded Optimization for Discovering Optimal Lateral Handling Bicycles}
\end{flushleft}

%%%%%%%%%%%%%%%% authors %%%%%%%%%%%%%%%
\begin{flushleft}
  {\fontsize{12}{14}{Jason K. Moore and Mont Hubbard}\\}
  \textit{Mechanical and Aerospace Engineering\\
          University of California, Davis\\
          One Shields Avenue, Davis, CA, USA 95817\\
          e-mail: jkm@ucdavis.edu, mhubbard@ucdavis.edu}
\end{flushleft}

% - Abstract
% - Introduction
% - Parameterization
% - Bounds and Constraints
% - Free Parameters
% - Optimization
% - Solutions
%   - Parameters
%   - Eigenvalues
%   - HQM
% - Discussion and Conclusion

\section*{Abstract}
%
Previously, we introduced a method of optimizing four primary geometric
parameters of a bicycle's design to minimize its Handling Quality Metric. Here
we expand that method to optimize 23 geometric and inertial parameters in the
benchmark parameterimization of the linear Whipple-Carvallo bicycle model. To
ensure physically realizable bicycle designs we include 7 equality constraints,
21 inequality constraints, and maximal and minamal bounds on each free
opmization parameter. This improves over the prior work by expanding the search
space with many more parameters and the guarntee of realizability. We close by
presenting five bicycle designs that have optimal later handling qualities. The
bicycles are not generally self-stable and have some ununusal characteristics.

\section{Introduction}
%
Physical design features of ground vehicles can affect their lateral handling
qualities. Geometry, mass, and mass distribution of the vehicle's components as
well as tire characteristics are primary contributors to poor and good handling
due to their important influence on the vehicle's dynamics. In past work, we
have presented a theoretical and computational framework for assessing the
lateral task-independent handling qualities of simplified single track vehicle
designs~\cite{Hess2012,Moore2012}. In subsequent work, we showed that
minimizing our proposed handling quality metric (HQM) can produce theoretically
optimal handling designs when only four geometric parameters are explored as
the optimization variables~\cite{Moore2016}. The present work's goal is to
expand this optimization problem to all of the geometry, mass, and inertial
parameters present in the linear Whipple-Carvallo bicycle
model~\cite{Meijaard2007}.  This broadens the search space considerably but we
constrain it so that only realizable optimal bicycle designs are discovered. To
do so, we formulate a constrained optimization problem and use derivative-free
optimization to discover optimal, yet realizable, bicycle designs.

\section{Bicycle Model Parameterization}
%
Our problem formulation relies on a new bicycle model parameterization that
reflects both a reformulation of and addition to the benchmark parameterization
of the linear Whipple-Carvallo bicycle model~\cite{Meijaard2007}. We call this
the ``principal parameterization'' as opposed to the ``benchmark
parameterization'' in~\cite{Meijaard2007}. This parameterization differs from
the benchmark parameterization in three ways. Firstly, the person and rear
frame are treated as separate rigid bodies each with their on inertial
parameters. Secondly, we express the inertial parameters of each rigid body in
terms of central principal radii of gyration to decouple the mass from the
inertia terms.  Lastly, we introduce two simple dimensional parameters that
define the geometric extents of the person which are used to constrain the
location of the person's body. Table~\ref{tab:parameters} provides the
parameters and the reference values which are derived from the measurements of
a Batavus Browser Bicycle and the rider ``Jason'' presented
in~\cite{Moore2012}. This parameterization can be transformed into the
benchmark parameterization readily, but not vice versa.
%
\begin{figure}
  \centering
  \includegraphics[width=4in]{../figures/reference-geometry.png}
  \caption{Depiction of the principal parameters of the Batavus Browser
    with rider ``Jason''. The solid black lines represent the essential bicycle
    geometry. The dotted black line represents the steer axis. The inertial
    properties of the five rigid bodies: front wheel (orange), rear wheel
    (purple), rear frame (blue), front frame (green), and person (red) are
    shown with the mass center and the extents of the centroidal principal
    radii of gyration of each rigid body. The primary principal angles,
    $\alpha_{D,P,H}$, are defined as the angle about the $y$ axis from $x$ to
    the maximum principal axes in the $XZ$ plane.}
\end{figure}
%
\begin{table}
  \caption{Full set of 47 principal parameters and their default values derived
    from the measurements in~\cite{Moore2012} of the Batavus Browser bicycle
    and rider ``Jason''. Note that the benchmark parameters are also used in
    the paper and are defined in~\cite{Meijaard2007}.}
  \label{tab:parameters}
  \small
  \centering
  \begin{tabular}{llll}
    \toprule
    Variable & Value & Units & Description \\
    \midrule
    % 5
    $c$       &  0.068581           & \si{\meter} & Trail \\
    $w$       &  1.1210             & \si{\meter} & Wheelbase \\
    $\lambda$ & 0.39968             & \si{\radian} & Steer axis tilt \\
    $g$       &  9.81               & \si{\meter\per\second\squared} & Acceleration due to gravity \\
    $v$       &  3.0, 5.0, 7.0, 9.0 & \si{\meter\per\second} & Forward speed \\
    Rear Wheel [R] & & \\
    \midrule
    % 8
    $m_R$     & 3.11 & \si{\kilogram}  & Mass \\
    $r_R$     & 0.34096 & \si{\meter}  & Radius \\
    $x_R$     & 0.0 & \si{\meter}      & $X$ mass center coordinate \\
    $y_R$     & 0.0 & \si{\meter}      & $Y$ mass center coordinate \\
    $z_R$     & -0.34096 & \si{\meter} & $Z$ mass center coordinate \\
    $k_{Raa}$ & 0.17050 & \si{\meter}  & Central principal radii of gyration about $A_R$ \\
    $k_{Rbb}$ & 0.17050 & \si{\meter}  & Central principal radii of gyration about $B_R$ \\
    $k_{Ryy}$ & 0.22136 & \si{\meter}  & Central principal radii of gyration about $Y$ \\
    Front Wheel [F] & & \\
    \midrule
    % 8
    $m_F$     & 2.02 & \si{\kilogram} & Mass \\
    $r_F$     & 0.34353 & \si{\meter} & Radius \\
    $x_F$     & 1.1210 & \si{\meter} & $X$ mass center coordinate \\
    $y_F$     & 0.0 & \si{\meter} & $Y$ mass center coordinate \\
    $z_F$     & -0.34353 & \si{\meter} & $Z$ mass center coordinate \\
    $k_{Faa}$ &  0.20917 & \si{\meter} & Central principal radii of gyration about $A_F$ \\
    $k_{Fbb}$ &  0.20917 & \si{\meter} & Central principal radii of gyration about $B_F$ \\
    $k_{Fyy}$ &  0.27179 & \si{\meter} & Central principal radii of gyration about $Y$ \\
    Person [P] & & \\
    \midrule
    % 10
    $l_P$      &  1.728   & \si{\meter} & Body length \\
    $w_P$      &  0.483   & \si{\meter} & Body width \\
    $m_P$      &  83.5    & \si{\kilogram} & Mass \\
    $x_P$      &  0.31577 & \si{\meter} & $X$ mass center coordinate \\
    $y_P$      & 0.0      & \si{\meter} & $Y$ mass center coordinate \\
    $z_P$      & -1.0990  & \si{\meter} & $Z$ mass center coordinate \\
    $k_{Paa}$  &  0.36797 & \si{\meter} & Central principal radii of gyration about $A_P$ \\
    $k_{Pbb}$  &  0.15276 & \si{\meter} & Central principal radii of gyration about $B_P$ \\
    $k_{Pyy}$  &  0.36717 & \si{\meter} & Central principal radii of gyration about $Y$ \\
    $\alpha_P$ & 0.18618 & \si{\radian} & Angle about $Y$ between $A_P$ and $X$ \\
    Front Frame [H] & & \\
    \midrule
    % 8
    $m_H$ & 3.2200 & \si{\kilogram}     & Mass \\
    $x_H$ & 0.86695 & \si{\meter}       & $X$ mass center coordinate \\
    $y_H$ & 0.0 & \si{\meter}           & $Y$ mass center coordinate \\
    $z_H$ & -0.74824 & \si{\meter}      & $Z$ mass center coordinate \\
    $k_{Haa}$ & 0.29556 & \si{\meter}   & Central principal radii of gyration about $A_H$ \\
    $k_{Hbb}$ & 0.14493 & \si{\meter}   & Central principal radii of gyration about $B_H$ \\
    $k_{Hyy}$ & 0.27630 & \si{\meter}   & Central principal radii of gyration about $Y$ \\
    $\alpha_H$ & 0.36995 & \si{\radian} & Angle about $Y$ between $A_H$ and $X$ \\
    Rear Frame [D] & & \\
    \midrule
    % 8
    $m_D$ &  9.8600 & \si{\kilogram}    & Mass \\
    $x_D$ &  0.27595 & \si{\meter}      & $X$ mass center coordinate \\
    $y_D$ & 0 & \si{\meter}             & $Y$ mass center coordinate \\
    $z_D$ & -0.53784 & \si{\meter}      & $Z$ mass center coordinate \\
    $k_{Daa}$ &  0.28587 & \si{\meter}  & Central principal radii of gyration about $A_D$ \\
    $k_{Dbb}$ &  0.22079 & \si{\meter}  & Central principal radii of gyration about $B_D$ \\
    $k_{Dyy}$ &  0.36539 & \si{\meter}  & Central principal radii of gyration about $Y$ \\
    $\alpha_D$ &  1.1722 & \si{\radian} & Angle about $Y$ between $A_D$ and $X$ \\
    \bottomrule
  \end{tabular}
\end{table}

\section{Bounds and Constraints}
%
The optimal principal parameters are subject to a set of constraints designed
to ensure that a physically realizable bicycle is obtained from the
optimization procedure. These constraints are made up of bounds on the free
parameters and both equality and inequality constraints among the parameters.
Below the basic constraint concepts presented and grouped by the associated
rigid body or collection thereof:
%
\begin{description}
  \item[Total $T$] The combination of the five rigid bodies.
    \begin{itemize}
      \item The likely physical extents of the rigid bodies must exist above
        the ground plane.
      \item Both bicycle and rider are symmetric about the rider's sagittal
        plane.
      \item The total mass is below a reasonably human lift-able amount.
      \item The wheels cannot overlap.
      \item The bicycle cannot topple forward during hard braking or backward
        during hard acceleration.
      \item The closed-loop path tracking controlled system~\cite{Hess2012}
        must be stable. This is required to obtain a valid HQM value.
    \end{itemize}

  \item[Rider $P$] A single rigid body represents the rider.
    \begin{itemize}
      \item Rider mass is that of an average person.
      \item The rider's joint angles are fixed in a nominal configuration
        typical of upright bicycling and the resulting mass distribution is
        derived from standard body segment estimation methods.
      \item The rider cannot penetrate the ground.
    \end{itemize}

  \item[Frames $H$, $D$] Front frame (handlebar + fork) and rear frame
    \begin{itemize}
      \item The rear frame is planar in nature and the front frame's moments of
        inertia are consistently dependent.
      \item The mass and inertia of the frames are positive and large enough to
        be constructed from a steel space frame.
    \end{itemize}

  \item[Wheels $F$, $R$] Both front and rear wheels have identical constraints.
    \begin{itemize}
      \item Wheel radius and mass must be positive and be greater than a
        minimum value.
      \item Wheels are inertially wheel-like, i.e. symmetric about each plane
        and most of the mass is at the rim.
    \end{itemize}

\end{description}

These bounds, equality, and inequality constraints are presented mathematically
in tables \ref{tab:bounds}, \ref{tab:equality-constraints},
\ref{tab:inequality-constraints} respectively and explained in more detail in
the following sections.
%
\begin{table}
  \caption{Parameter lower and upper bounds.}
  \label{tab:bounds}
  \centering
  \begin{tabular}{ccccc}
    \toprule
    Min & & Parameter & & Max\\
    \midrule
    $-\infty$ & $\leq$ & $w       $  &  $\leq$  &  $\infty$ \\
    $-\infty$ & $\leq$ & $c       $  &  $\leq$  &  $\infty$ \\
    $-\pi/2 $ & $\leq$ & $\lambda $  &  $\leq$  &  $\pi/2$ \\
    $1.0\si{\kilogram}$ & $\leq$ & $m_D     $  &  $\leq$  &  $\infty$ \\
    $-\infty$ & $\leq$ & $x_D     $  &  $\leq$  &  $\infty$ \\
    $-\infty$ & $\leq$ & $z_D     $  &  $\leq$  &  $0.0$    \\
    $0.0    $ & $\leq$ & $k_{Daa} $  &  $\leq$  &  $\infty$ \\
    $0.0    $ & $\leq$ & $k_{Dbb} $  &  $\leq$  &  $\infty$ \\
    $-\pi/2 $ & $\leq$ & $\alpha_D$  &  $\leq$  &  $\pi/2$  \\
    $-\infty$ & $\leq$ & $x_P     $  &  $\leq$  &  $\infty$ \\
    $-\infty$ & $\leq$ & $z_P     $  &  $\leq$  &  $0.0$    \\
    $-\pi/2 $ & $\leq$ & $\alpha_P$  &  $\leq$  &  $\pi/2$ \\
    $0.25   $ & $\leq$ & $m_H     $  &  $\leq$  &  $\infty$ \\
    $-\infty$ & $\leq$ & $x_H     $  &  $\leq$  &  $\infty$ \\
    $-\infty$ & $\leq$ & $z_H     $  &  $\leq$  &  $0.0$ \\
    $0.0    $ & $\leq$ & $k_{Haa} $  &  $\leq$  &  $\infty$ \\
    $0.0    $ & $\leq$ & $k_{Hbb} $  &  $\leq$  &  $\infty$ \\
    $0.0    $ & $\leq$ & $k_{Hyy} $  &  $\leq$  &  $\infty$ \\
    $-\pi/2 $ & $\leq$ & $\alpha_H$  &  $\leq$  &  $\pi/2$ \\
    $0.127\si{\meter}$ & $\leq$ & $r_R     $  &  $\leq$  &  $\infty$ \\
    $1.0\si{\kilogram}$ & $\leq$ & $m_R     $  &  $\leq$  &  $\infty$ \\
    $0.127\si{\meter}$ & $\leq$ & $r_F     $  &  $\leq$  &  $\infty$ \\
    $1.0\si{\kilogram}$ & $\leq$ & $m_F     $  &  $\leq$  &  $\infty$ \\
    \bottomrule
  \end{tabular}
\end{table}
%
\begin{table}
  \caption{Equality constraints.}
  \label{tab:equality-constraints}
  \centering
  \begin{tabular}{lll}
    \toprule
    Constraint & Equation & Description \\
    \midrule
    $g_1$ & $I_{Dyy} = \sqrt{I_{Dxx}^2 + I_{Dzz}^2}$ & Rear frame is planar. \\
    $g_2$ & $k_{Ryy} = r_R$  & Rear wheel is a ring \\
    $g_3$ & $k_{Raa} = k_{Ryy}/2$ & Rear wheel is a ring \\
    $g_4$ & $k_{Rbb} = k_{Ryy}/2$ & Rear wheel is a ring \\
    $g_5$ & $k_{Fyy} = r_F$  & Front wheel is a ring \\
    $g_6$ & $k_{Faa} = k_{Fyy}/2$ & Front wheel is a ring \\
    $g_7$ & $k_{Fbb} = k_{Fyy}/2$ & Front wheel is a ring \\
    \bottomrule
  \end{tabular}
\end{table}
%
\begin{table}
  \caption{Inequality constraints.}
  \label{tab:inequality-constraints}
  \centering
  \begin{tabular}{lll}
    \toprule
    Constraint & Equation & Description \\
    \midrule
    $c_1$    & $\sqrt{I_{Hxx}^2+I_{Hzz}^2} \geq I_{Hyy}$           & Consistent moments of inertia. \\
    $c_2$    & $0 \geq z_P + l_P/2 \cos{\alpha_P}$                 & Person cannot penetrate ground. \\
    $c_3$    & $0 \geq z_P + w_P/2 \sin{\alpha_P}$                 & Person cannot penetrate ground. \\
    $c_4$    & $0 \geq z_P - l_P/2 \cos{\alpha_P}$                 & Person cannot penetrate ground. \\
    $c_5$    & $0 \geq z_P - w_P/2 \sin{\alpha_P}$                 & Person cannot penetrate ground. \\
    $c_6$    & $x_T \geq |z_T|/4$                                  & Maximum acceleration of $1/4g$. \\
    $c_7$    & $w-x_T \geq 3/4|z_T|$                               & Maximum deceleration of $3/4g$. \\
    $c_8$    & $2k_{Hyy} \geq \sqrt{(x_H - w)^2 + (z_H + r_F)^2}$  & Minimal inertial spread. \\
    $c_9$    & $2k_{Dyy} \geq \sqrt{(x_D - 0)^2 + (z_D + r_R)^2}$  & Minimal inertial spread. \\
    $c_{10}$ & $w \geq r_F +r_R$                                   & Non-overlapping wheels. \\
    $c_{11}$ & $25\si{\kg} \geq m_D + m_H + m_R + m_F$             & Maximum bicycle mass. \\
    $c_{12}$ & $-z_D \geq 1.4 k_{Dyy}$                             & Rear frame cannot penetrate ground. \\
    $c_{13}$ & $-z_H \geq 1.4 k_{Hyy}$                             & Front frame cannot penetrate ground. \\
    $c_{14,\ldots,21}$ & $0 \geq s_1,\ldots,s_8$                   & Closed loop stability. \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Person [P]}
%
We assume that the person's joint configurations are such that they are in a
nominal configuration for pedaling, i.e. an average normal everyday riding
position on a typical bicycle. We retain the same configuration as they were
seated on the Batavus Browser bicycle. The person is assumed to be symmetric
about the $XZ$ plane. We allow the rider to be rotated about the $Y$ axis and
positioned anywhere within the plane of symmetry above the ground.

To prevent the rider from being positioned and oriented such that their body is
not penetrating the ground we introduce two dimensions that define a cross
whose apex is at the center of mass of the person and the cross axes are
parallel to the principal axes in the $XZ$ plane. $l_P / 2$ is the distance
along the principal axis to the tip of the toes and $w_P / 2$ is the distance
along the second principal axes to the tip of the hands. The constraints
$c_2,\ldots,c_5$ are derived from these rules.

\subsection{Front Frame [H]}
%
The front frame is symmetric about the $XZ$ plane so $I_{Hxy}, I_{Hyz} = 0$. We
allow for any angular orientation of the $XZ$ principal directions but limit
the angle to $-\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2}$. We prevent the
rear frame from penetrating the ground by limiting the inertial spread with
respect to its mass center, $c_{13}$, but also set a minimum inertial spread to
ensure a frame can span from the rear wheel to the mass center of the rear
frame, $c_8$. The spread factor in $c_{13}$ of 1.4 is based on the ratio of
geometrical spread of a typical bicycle frame and its radius of gyration. The
front frame is not planar but is narrow with respect to the $XZ$ plane, which
is enforced by constraint $c_1$.

\subsection{Rear Frame [D]}
%
Several constraints are set for the rear frame. We constrain the rear frame to
be planar, $g_1$, and symmetric with respect to the $XZ$ plane. We prevent the
rear frame from penetrating the ground by limiting the inertial spread with
respect to its mass center, $c_{12}$, but also set a minimum inertial spread to
ensure a frame can span from the rear wheel to the mass center of the rear
frame, $c_9$. The spread factor in $c_{12}$ of 1.4 is based on the ratio of
geometrical spread of a typical bicycle frame and its radius of gyration.
Several parameters are bounded. We require the rear frame mass to be positive,
the center of mass not penetrate the ground, and we allow for any angular
orientation of the $XZ$ principal directions but limit the angle to
$-\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2}$ as angle beyond that are
redundant.

\subsection{Front [F] and Rear [R] Wheels}
%
We enforce the assumption that both wheels have moments of inertia of that of a
simple ring, $g_2 \ldots g_7$ and that mass and radius should be greater than a
minimal size based on small purchasable spoked wheel with tire.

\subsection{Total Bike [T]}
%
The trail and wheelbase can take on any real values. The steer axis tilt is
limited to $\pm90$ degrees. We introduce a constraint $c_{10}$ that prevents
the wheels from physically overlapping and require that the bicycle be
lift-able by an average person, $c_{11}$. Finally, we require that the bicycle
not topple forward during hard breaking or backward during hard acceleration
with:

\begin{align}
  -\frac{3g}{4} < \textrm{acceleration} < \frac{g}{4}\textrm{.}
\end{align}

This translates to two constraints, $c_6,c_7$ that bound the total center of
mass $(x_T,z_T)$ in a triangle in the $XZ$ plane. Lastly, we constrain the
eight closed loop eigenvalues associated with the controller in~\cite{Hess2012}
to be stable, i.e. have negative real parts. These are expressed in
constraints $c_{14},\ldots,c_{21}$. Closed loop stability is required for the
HQM to provide a meaningful result.

\section{Optimization}
%
The above constraints leave 23 of the 47 parameters free for optimizing which
we collect in the vector $\mathbf{p}\in\mathbb{R}^{23}$ and define as:
%
\begin{align}
  \begin{split}
    \mathbf{p} = [ &
       w        \quad
       c        \quad
       \lambda  \quad
       m_D      \quad
       x_D      \quad
       z_D      \quad
       k_{Daa}  \quad
       k_{Dbb}  \quad
       \alpha_D \quad
       x_P      \quad
       z_P      \quad
       \alpha_P \\
     & m_H      \quad
       x_H      \quad
       z_H      \quad
       k_{Haa}  \quad
       k_{Hbb}  \quad
       k_{Hyy}  \quad
       \alpha_H \quad
       r_R      \quad
       m_R      \quad
       r_F      \quad
       m_F]
  \end{split}
\end{align}

Our objective in the optimization is to minimize the peak HQM value subject to
the bounds, $\mathbf{p}^L,\mathbf{p}^U$, and the constraints
$\mathbf{g}(\mathbf{p}),\mathbf{c}(\mathbf{p})$ . Given a set of bicycle model
parameter values we generate a bandwidth limited human-like controller using
the methods in ~\cite{Moore2012}. Once the closed loop stable controller is
constructed, the HQM can be computed as per the definition in \cite{Hess2012}
and the scaler peak value returned as the objective $J$. This problem is
presented as a non-linear programming problem in the following equation.
%
\begin{equation}
  \begin{aligned}
    & \underset{\mathbf{p}}{\text{minimize}} & & J(\mathbf{p})=max(\textrm{HQM}(\mathbf{p})) \\
    & \text{subject to} & & \\
    & & & \mathbf{g}(\mathbf{p}) \leq \mathbf{0} \\
    & & & \mathbf{c}(\mathbf{p}) = \mathbf{0} \\
    & & & \mathbf{p}^L \leq \mathbf{p} \leq \mathbf{p}^U
  \end{aligned}
\end{equation}

We make use of the derivative-free optimizer CMA-ES~\cite{Hansen1996} to find
solutions to this problem. The optimization supports parameter bounds and
equality constraints but does not support inequality constraints. To get around
this limitation we move the inequality constraints into the objective function
and penalize the objective if the constraints are violated with the following
rules:
%
\begin{align}
  J(\mathbf{p}) =
  \begin{cases}
    max(\textrm{HQM}(\mathbf{p})) & \textrm{if} \quad all(\mathbf{g}(\mathbf{p})) \leq 0 \\
    30 + ||\mathbf{g}_{+}(\mathbf{p})||/10 & \textrm{if} \quad
      any(\mathbf{g}(\mathbf{p})) > 0 \textrm{ and } ||\mathbf{g}_{+}(\mathbf{p})|| < 30 \\
    ||\mathbf{g}_{+}(\mathbf{p})|| & \textrm{if} \quad
      any(\mathbf{g}(\mathbf{p})) > 0 \textrm{ and } ||\mathbf{g}_{+}(\mathbf{p})|| \geq 30
  \end{cases}
\end{align}
where $||\mathbf{g}_{+}||$ is the norm of the vector of positive elements of
$\mathbf{g}$ and $any,all$ are ``any elements of'' and ``all elements of'',
respectively.

This creates a discontinuous objective function but in practice the CMA-ES
algorithm is able to move into the parameter space where all the constraints
are satisfied and find a (local) minima. For our purposes, this sufficiently
finds parameter values that produce optimally handling bicycle designs.

\section{Results}
%
We discover four bicycles for different design speeds (3, 5, 7, and
9~\si{\meter\per\second}) that have an optimally low HQM, see
Table~\ref{tab:hqm}, and satisfy all constraints and parameter bounds. The
optimal parameters for these four bicycles are presented in
Table~\ref{tab:optimal-values}.
%
\begin{table}
  \caption{Peak HQM values for the reference bicycle and the optimal bicycles
    at each design speed.}
  \label{tab:hqm}
  \centering
  \begin{tabular}{llll}
    \toprule
    Speed [m/s] & Reference Peak HQM & Optimal Peak HQM & Percent Improvement \\
    \midrule
    3 & 13.075 & 2.012 & 85\% \\
    5 & 4.521  & 0.012 & 100\% \\
    7 & 3.043  & 0.022 & 99\% \\
    9 & 2.338  & 0.839 & 64\% \\
    \bottomrule
  \end{tabular}
\end{table}
%
\begin{table}
  \caption{Optimal principal parameter values for each design speed. Values
    with an asterisk are at a bound.}
  \label{tab:optimal-values}
  \centering
  \begin{tabular}{lSSSS}
    \toprule
    $v$ & 3~\si{\meter\per\second} & 5~\si{\meter\per\second} & 7~\si{\meter\per\second} & 9~\si{\meter\per\second} \\
    \midrule
    $c$       & 0.688  & -0.005 & 0.001  & -0.484 \\
    $w$       & 2.866  & 0.847  & 1.157  & 5.557  \\
    $\lambda$ & -0.213 & 0.271  & -0.028 & 0.239  \\
    \midrule
    $m_D$ & 3.22   &  11.53 & 10.32  & 7.02 \\
    $x_D$ & 0.958  &  0.287 &  0.427 & 1.020 \\
    $z_D$ & -2.605 & -2.719 & -2.976 & -4.065 \\
    $\alpha_D$ & 0.663  & 0.915  & 1.123  & 1.077 \\
    % Should aa be larger than bb?
    $k_{Daa}$  & 1.449  & 1.218  &  1.324 & 0.861 \\
    $k_{Dbb}$  & 0.724  & 1.387  &  1.290 & 2.014 \\
    $k_{Dyy}$  & 1.471  & 1.559  &  1.555 & 2.031 \\
    \midrule
    $m_H$  & 0.25*  & 0.25* &    0.49 &   3.54 \\
    $x_H$     & 2.356   & 0.532 &    1.145 &    5.615 \\
    $z_H$     & -1.298  & -0.781 &   -1.340 &   -1.776 \\
    $\alpha_H$ & 0.572  & -1.265 & 1.571  & 0.238 \\
    $k_{Haa}$ & 0.186  & 0.0491 &   0.000* &       2.846 \\
    $k_{Hbb}$ & 1.259   & 1.145 &     2.006 &    0.168 \\
    $k_{Hyy}$ & 0.636  & 0.383 &    0.818 &   0.669 \\
    \midrule
    $m_F$     & 1.62  & 4.40  & 2.64  & 5.51  \\
    $r_F$     & 0.710 & 0.252 & 0.127* & 2.063 \\
    $k_{Faa}$ & 0.355 & 0.126 & 0.064 & 1.031 \\
    $k_{Fyy}$ & 0.710 & 0.252 & 0.127 & 2.063 \\
    \midrule
    $x_P$      & 0.765  & 0.276  &  0.526 & 1.079 \\
    $z_P$      & -3.194 & -0.453 & -0.586 &-2.662 \\
    $\alpha_P$ & -0.661 & 1.150  & -0.836 & 1.558 \\
    \midrule
    $m_R$   & 12.63  & 8.36  &  6.31 & 2.25 \\
    $r_R$   & 0.958  & 0.506 & 0.146 & 3.027 \\
    $k_{Raa}$ & 0.479  & 0.253 & 0.073 & 1.514 \\
    $k_{Ryy}$ & 0.958  & 0.506 & 0.146 & 3.027 \\
    \bottomrule
  \end{tabular}
\end{table}

The pictorial representation of the bicycles are presented in
Figure~\ref{fig:optimal-geometries}. The most striking feature is that all of
the bicycle are larger that the reference bicycle in some way.
%
\begin{figure}
  \centering
  \includegraphics[width=6in]{../figures/optimal-geometries.png}
  \label{fig:optimal-geometries}
  \caption{Depictions of the bicycle geometry and geometric representations of
    the inertial quantities for the reference bicycle and four optimal solutions
    at 3, 5, 7, and 9 m/s. Five rigid bodies are shown for each bicycle: front
    wheel (orange), rear wheel (purple), rear frame (blue), front frame
    (green), and person (red). The solid black lines represent the essential
    bicycle geometry. The dotted black line represents the steer axis. The
    solid colored curves represent the contours of solid ellipsoids with
    equivalent inertia as the principal inertia of the associated rigid body.
    The dotted colored lines represent the extents of the centroidal radii of
    gyration of each rigid body.}
\end{figure}

We
belive these bicycles to be physically realizable with minor differences.

  The 3 m/s bicycle places both the rider's mass center and
the rear frame's mass center close to 3 meters above the ground plane. The 5
and 7 m/s bicycles have wheelbase values that are simlar to the reference
bicycle, but 3 and 9 m/s have very large wheel bases. The 9 m/s bicycle is very
large overall. Only the 5 m/s bicycle is of a scale close to a typical bicycle.
It is noteable that the 9 m/s bicycle has signficant negative trail.

3
- Rider is over 3 m in the air and leaned forward.
- 3 m wheel base with 0.7 meter positive trail
- wheels are over double the diameter of browser wheels
- minimum front frame mass

5 m/s
- rider is laid back a 0.5 meter off ground
- < 1 m wheelbase
- essentially zero trail (5 mm negative)
- minimum front frame mass

7 m/s
- miminal rotatoinal inertia about the steer axis
- tall slender fork/handlebar
- no trail (1 mm postive)
- half sized wheels (~12 inches)
- normal wheelbase (1.157)
- rider tipped forward and low (0.6 m)
- vertical steer axis
- 3 m high frame center of mass

9
- very large 5.6 m wheelbase
%
\begin{figure}
  \centering
  \includegraphics[width=6in]{../figures/optimal-eigenvalues.png}
  \label{fig:optimal-eigenvalues}
  \caption{}
\end{figure}

\section{Discussion and Conclusion}
%
- need constraints to keep rider out of wheels
- need more realistic constraints for the frame mass and inertia based on space
frame
\section{Reproducibility}
%
All of the source code, data, and documents needed to reproduce the presented
results and this paper can be found at the repository hosted at
\url{https://github.com/moorepants/BMD2019}.

\bibliographystyle{acm}
\bibliography{BMD2019}

\end{document}